Generating Images From Captions\\ With Attention

### 2015-11-01 iclr-paper-new.tex
a

To generate an image $\oimage$ conditioned on the caption information $\bf{y}$, 
we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ timesteps.

Unlike the original DRAW network where latent variables are independent 
spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.  
Namely, the mean and variance of the prior distribution over $\Lat_t$ are parameterized by: 
\begin{align}
\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
\mu_{t}(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\\
\sigma_{t}(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big),
\end{align}
where $W_{\mu} \in \mathbb{R}^{D \times n}$, $W_{\sigma} \in \mathbb{R}^{D \times n}$ are 
the learnt model parameters, where $n$ is the dimensionality of $\hdec_{t}$,
the hidden state of the generative LSTM. 
Similar to \cite{bachman_sdm}, we have observed that 
the model performance is improved by including dependencies between latent variables.

Formally, an image is generated by iteratively computing the following set of 
equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
\begin{align}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
s_{t} &= \aalign(\hdec_{t-1}, \hlangall), \\
\hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
\canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \sigmoid(\canv_{T,i}),
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
where $\decoder$
is the function defined by the LSTM network with \textit{forget} gates 
\citep{gers_forget} at a single time-step. To generate the next 
hidden state $\hdec_t$, 
the decoder $\decoder$ takes its previous hidden state $\hdec_{t-1}$ and
combines it with the input from both the sampled latent variable $z_t$ and 
caption representation $s_t$ at time-step $t$, where
$[v_1, v_2]$ denotes the concatenation of the vectors $v_1$ and $v_2$ into the single vector.
The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, with learnt bias $b$.

The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. 
Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities 
$\alpha_{1...N}^{t}$:
\beqa
s_t=\aalign(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
\eeqa
The corresponding alignment probabilities $\alpha_{1...N}^{t}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$ 
and then normalizing them: %, as in \citep{bahdanau_mt}:
\beqa
\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}
\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t} + b)}, 
\eeqa
where $U,W,b$ are the learned model parameters.

%is a standard \textit{Long Short-Term Memory} architecture \citep{hochreiter_lstm} with \textit{forget} gates \citep{gers_forget}, $[v_1, v_2]$ is the concatenation of the vectors $v_1$ and $v_2$ into the single vector, and $\canv_0$, $\hdec_0$ are initialized to the the learned biases.

The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas $c_t$ with smoothly varying location and zoom (same as in \citep{gregor_draw}), with 
$c_0=0$. 
Finally, the $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,

@@

The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}. The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training. 
In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular timestep while completely ignoring the other digit. 
Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process. However in both cases, the model placed small attention values on the words describing the position of digits in the images.

@@

We also tried to analyze the way the model generated images. Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.

>>>>

To generate an image $\oimage$ conditioned on the caption information $\bf{y}$, 
we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ time-steps.

Unlike the original DRAW network where latent variables are independent 
spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.  
Namely, the mean and variance of the prior distribution over $\Lat_t$ are parameterized by: 
\begin{align}
\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
\mu_{t}(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\\
\sigma_{t}(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big),
\end{align}
where $W_{\mu} \in \mathbb{R}^{D \times n}$, $W_{\sigma} \in \mathbb{R}^{D \times n}$ are 
the learned model parameters, and $n$ is the dimensionality of $\hdec_{t}$,
the hidden state of the generative LSTM. 
Similar to \cite{bachman_sdm}, we have observed that 
the model performance is improved by including dependencies between latent variables.

Formally, an image is generated by iteratively computing the following set of 
equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
\begin{align}
\label{eq:x_hat}
\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
s_{t} &= align(\hdec_{t-1}, \hlangall), \\
\label{eq:decoder}
\hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
\label{eq:writeop}
\canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
\label{eq:write}
%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}
The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps~\citep{bahdanau_mt}.
Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities
$\alpha_{1...N}^{t}$:
\beqa
s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
\eeqa
The corresponding alignment probabilities $\alpha_{1...N}^{t}$ are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t-1}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
and then normalizing them: %, as in \citep{bahdanau_mt}:
\beqa
\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)}{\sum_{j=1}^{N}
\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t-1} + b)},
\eeqa
where $U \in \mathbb{R}^{d \times m}$, $W \in \mathbb{R}^{d \times n}$ and 
$b \in \mathbb{R}^{d \times 1} $ are the learned model parameters of the alignment model.

The $\decoder$ function of \Eqref{eq:decoder} 
is defined by the LSTM network with \textit{forget} gates 
\citep{gers_forget} at a single time-step. To generate the next 
hidden state $\hdec_t$, 
the decoder $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
combines it with the input from both the latent sample $z_t$ and 
sentence representation $s_t$, where
$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, 
with learned bias $b$.

The output of the decoder $\hdec_t$ is then passed through the $\writeop$ operator 
which is added to a cumulative canvas matrix $c_t$, with $c_0=0$ (\Eqref{eq:writeop}). 
The $\writeop$ operator
applies an array of Gaussian filters on the patch, placing 
it on the canvas $c_t$ with smoothly varying location and zoom (same as 
defined in \citep{gregor_draw}).

Finally, each pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli 
distribution with the mean vector $\sigmoid(\canv_{T})$ over the image pixels 
$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$,

@@

The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}. The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training. 
In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular time-step while completely ignoring the other digit. 
Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process. However in both cases, the model placed small attention values on the words describing the position of digits in the images.

@@

We also tried to analyze the way the model generated images. Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular time-steps.
